---
output: hugodown::md_document
title: "TidyTuesday Penguins"
summary: "Penguins, Datapasta, and a Github Cawebsite"
author: "rbkeeney"
tags: []
date: 2020-07-28
---

## Introduction 

In my 1st ever TidyTuesday, I felt inspired by Asmae Toumi's (@asmae_toumi) recent twitter post about creating a caladown site through github. Here's a link to the guide she posted: [Youtube Video](https://www.youtube.com/watch?v=HtQhG80MKQE)

It took Asmae 15 minutes to deploy her site.

*It took me about 6 hours.* I've never used git or created a github repo from R before today. A summary of how it went?

* Gownload git
* Google how to use git
* Learn a new cuss words
* Crash R a feww times
* Learn more cuss words
* Learn how to set directories and remove ".git/index.lock" files
* Create a SSH key
* Create a Github token
* Troubleshoot why R crashes when I run "commit" or "push"
* Fix it
* Repo is working!
* Break post code
* Ugh.

So. Much. Fun.

*and yes, eventually I got it figured out*

## Getting Started

Finally getting to the R code (the easy portion!), my Tidytuesday entry about creating a caladown site, using rMarkdown and some datapasta, and eventually making a simple animated plot.

First, let's install the our libraries and load them, for more detail check out the post  [here](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-07-28/readme.md)

```{r install it, message=FALSE, warning=FALSE}
#install.packages("tidytuesdayR")
library(tidytuesdayR)
library(tidyverse)
library(gganimate)
library(glue)
```

TidyTuesday makes is easy to get the data. BTW, ctrl+alt+i -> "insert new code section" is wonderful. For more on rMarkdown: [rMarkdown Intro](https://rmarkdown.rstudio.com/articles_intro.html)

```{r load it, message=FALSE, warning=FALSE}
tuesdata <- tidytuesdayR::tt_load('2020-07-28')
penguins <- tuesdata$penguins
```

## Datapasta

I also wanted to try out a neat little package out there called datapasta... you can load it like so if you don't have it. For more detail, you can visit [Datapasta](https://github.com/MilesMcBain/datapasta).

```{r install it again}
#install.packages("datapasta")
```

Datapasta allows you to paste data from the web directly into your code in about 2 seconds.

1. Control+C (windows)
2. Control+Shift+T (after setting up the new keyboard shortcuts)

Add in a little gt:gt() or knitr:kable() and you've got yourself a nice little table to play with, *additional formatting purely optional*. After the learning git + github all day, I'm going to use knitr:kable() can call it day.

```{r past it}
tibble::tribble(
               ~variable,    ~class,                                               ~description,
               "species", "integer",              "Penguin species (Adelie, Gentoo, Chinstrap)",
                "island", "integer",         "Island where recorded (Biscoe, Dream, Torgersen)",
        "bill_length_mm",  "double", "Bill length in millimeters (also known as culmen length)",
         "bill_depth_mm",  "double",   "Bill depth in millimeters (also known as culmen depth)",
     "flipper_length_mm", "integer",                                     "Flipper length in mm",
           "body_mass_g", "integer",                                       "Body mass in grams",
                   "sex", "integer",                                        "sex of the animal",
                  "year", "integer",                                            "year recorded"
     ) %>% 
     knitr::kable()
```

## Animated plot

Now I'm going to make a simple plot... flipper length x bill length should do nicely.

```{r plot it, warning=FALSE}
p1 <- penguins %>% 
     ggplot(aes(y = flipper_length_mm, x = bill_length_mm,color = species,group = species)) +
     geom_smooth(method=lm) +
     geom_point(size = 2, alpha = 0.6) +
     theme_bw()  +
     labs(title = "Palmer Penguins", 
          caption = "palmerpenguins data by Dr. Kristen Gorman",
          x = "Bill Length (mm)", 
          y = "Flipper Length (mm)")
```

Let's see what the plot looks like:

```{r print it, message=FALSE, warning=FALSE}
p1
```

To create an animation from the picture, the gganimate library is needed (loaded previously). I also like to use "glue" to easily paste information from the data. Here, I use the "closest_state" command to update the subtitle with which species is being displayed.

```{r prep it, warning=FALSE}

p1_anim <- p1 +
     transition_states(species) +
     labs(subtitle = "Species: {closest_state}") +
     ease_aes('cubic-in-out')
```

To display the animation, just call animate(). This take forever to knit, so I'm going to exclude it from the rMarkdown.

```{r move it, echo=FALSE, message=FALSE, warning=FALSE}
animate(p1_anim)
```

Finally, if you wanted to save the simple animation locally for later use (Twitter anyone?), you can save the last animated file using anim_save()

```{r save it, eval=FALSE, include=FALSE}
# anim_save(filename = 'Rplot_animation.gif', animation = last_animation(),path = "C:\\Users\\keene\\Desktop")
```
